---
title: "Streamflow Local Regression Estimation"
output: html_notebook
---

## Monthly Streamflow Kernel Density Estimation  

Split monthly data
```{r}
splitMnth <- split(gsMNF95, gsMNF95$Month)
```

Set current month: MARCH
```{r}
mnthData <- splitMnth$`03`
```

### Model fit and diagnostics

Fit local linear regression with different bandwidth alphas
```{r}
library(locfit)
par(mfrow=c(2,2))
par(mar=c(4,5,2,2))
alp <- c(0.2,0.4,0.6,0.8)
for(a in alp) {
  fit <- locfit(Flow~Lag1, data=mnthData, alpha=a, deg=1)
  plot(fit, get.data=T, col="blue",
       main=paste("alpha =",a), 
       xlab="Yi-1 (lag1) flow", 
       ylab="Yi flow")
}
```

### Local Regression Diagnostics and Goodness of Fit  

Smoothed Residual Plots: residuals vs. predictor to detect lack of fit
```{r}
par(mfrow=c(2,2))
par(mar=c(4,5,2,2))
alp <- c(0.2,0.4,0.6,0.8)
for(a in alp) {
  fit <- locfit(Flow~Lag1, data=mnthData, alpha=a, deg=2)
  res <- residuals(fit)
  fit2 <- locfit.raw(mnthData$Lag1, res, alpha=0.3)
  plot(mnthData$Lag1, res, pch=16, col="blue", cex=.5,
       main=paste("alpha =",a), 
       ylim=c(-10,10), 
       xlab="predictors", 
       ylab="residuals")
  lines(fit2)
  abline(h=0,lty=1)
}
```

Compare smoothing parameters to GCV score
```{r}
y <- mnthData$Flow
x <- mnthData$Lag1

gcvMnth <- gcvplot(y~x, kern="gauss", deg=1, ev=dat(), alpha=seq(0.01, 0.99, by=.01))
plot(gcvMnth,col="blue")

data.frame(bandwidth = gcvMnth$alpha, GCV = round(gcvMnth$values), DF=round(gcvMnth$df,1))
alphaMnth = gcvMnth$alpha[which.min(gcvMnth$values)]
alphaMnth
```

### Scatterplot Local Regression and k-NN Stochastic Simulation

K Nearest Neighbors of Lag-1 Month Streamflow
```{r}
library('knitr')
library(FNN)
xMnthLag <- mnthMean$mnthMean[2]
knnMnthLag <- get.knnx(data=mnthData$Lag1, query=xMnthLag, k=10, algorithm=c("kd_tree", "cover_tree", "CR", "brute"))
knnMnthLagPts <- mnthData$Lag1[knnMnthLag$nn.index]
knnMnthPts <- mnthData$Flow[knnMnthLag$nn.index]
```

Local regression and k nearest neighbors plot
```{r}
plotLocfit <- ggplot(data=mnthData, aes(x=Lag1, y=Flow)) +
  geom_point(colour="blue",shape=1) +
  geom_point(data=mnthData[knnMnthLag$nn.index,], aes(x=Lag1, y=Flow), colour="red", size=3, shape=2) +
  geom_smooth(method=locfit, formula = y ~ lp(x, nn=.35, deg=1),se = F)  +
  xlab(expression(month_Y-1~flow~"("*m^{3}*"/s)")) + 
  ylab(expression(month_Y~flow~"("*m^{3}*"/s)")) + 
  labs(title = "Local Regression", subtitle = "Month-on-Month Flow Dependency")
plotLocfit
```

---

### Monthly Local Regression Fits

JANUARY
```{r}
mnthData <- splitMnth$`01`
y <- mnthData$Flow
x <- mnthData$Lag1
fitJAN <- locfit(y~x, data=mnthData, alpha=0.3, deg=1)
resJAN <- residuals(fitJAN)
plot(fitJAN, get.data=T, xlab="mnth Yi-1 flow", ylab="mnth Yi flow")
```

FEBRUARY
```{r}
mnthData <- splitMnth$`02`
y <- mnthData$Flow
x <- mnthData$Lag1
fitFEB <- locfit(y~x, data=mnthData, alpha=0.4, deg=1)
resFEB <- residuals(fitFEB)
plot(fitFEB, get.data=T, xlab="mnth Yi-1 flow", ylab="mnth Yi flow")
```

MARCH
```{r}
mnthData <- splitMnth$`03`
y <- mnthData$Flow
x <- mnthData$Lag1
fitMAR <- locfit(y~x, data=mnthData, alpha=0.35, deg=1)
resMAR <- residuals(fitMAR)
plot(fitMAR, get.data=T, xlab="mnth Yi-1 flow", ylab="mnth Yi flow")
```

APRIL
```{r}
mnthData <- splitMnth$`04`
y <- mnthData$Flow
x <- mnthData$Lag1
fitAPR <- locfit(y~x, data=mnthData, alpha=0.35, deg=1)
resAPR <- residuals(fitAPR)
plot(fitAPR, get.data=T, xlab="mnth Yi-1 flow", ylab="mnth Yi flow")
```

MAY
```{r}
mnthData <- splitMnth$`05`
y <- mnthData$Flow
x <- mnthData$Lag1
fitMAY <- locfit(y~x, data=mnthData, alpha=0.3, deg=1)
resMAY <- residuals(fitMAY)
plot(fitMAY, get.data=T, xlab="mnth Yi-1 flow", ylab="mnth Yi flow")
```

JUNE
```{r}
mnthData <- splitMnth$`06`
y <- mnthData$Flow
x <- mnthData$Lag1
fitJUN <- locfit(y~x, data=mnthData, alpha=0.35, deg=1)
resJUN <- residuals(fitJUN)
plot(fitJUN, get.data=T, xlab="mnth Yi-1 flow", ylab="mnth Yi flow")
```

JULY
```{r}
mnthData <- splitMnth$`07`
y <- mnthData$Flow
x <- mnthData$Lag1
fitJUL <- locfit(y~x, data=mnthData, alpha=0.27, deg=1)
resJUL <- residuals(fitJUL)
plot(fitJUL, get.data=T, xlab="mnth Yi-1 flow", ylab="mnth Yi flow")
```

AUGUST
```{r}
mnthData <- splitMnth$`08`
y <- mnthData$Flow
x <- mnthData$Lag1
fitAUG <- locfit(y~x, data=mnthData, alpha=0.4, deg=1)
resAUG <- residuals(fitAUG)
plot(fitAUG, get.data=T, xlab="mnth Yi-1 flow", ylab="mnth Yi flow")
```

```{r}
# set month SEPTEMBER
mnthData <- splitMnth$`09`
y <- mnthData$Flow
x <- mnthData$Lag1
fitSEP <- locfit(y~x, data=mnthData, alpha=0.28, deg=1)
resSEP <- residuals(fitSEP)
plot(fitSEP, get.data=T, xlab="mnth Yi-1 flow", ylab="mnth Yi flow")
```

OCTOBER
```{r}
mnthData <- splitMnth$`10`
y <- mnthData$Flow[-1]
x <- mnthData$Lag1[-1]
fitOCT <- locfit(y~x, data=mnthData, alpha=0.3, deg=1)
resOCT <- residuals(fitOCT)
plot(fitOCT, get.data=T, xlab="mnth Yi-1 flow", ylab="mnth Yi flow")
```

NOVEMBER
```{r}
mnthData <- splitMnth$`11`
y <- mnthData$Flow
x <- mnthData$Lag1
fitNOV <- locfit(y~x, data=mnthData, alpha=0.44, deg=1)
resNOV <- residuals(fitNOV)
plot(fitNOV, get.data=T, xlab="mnth Yi-1 flow", ylab="mnth Yi flow")
```

DECEMBER
```{r}
mnthData <- splitMnth$`12`
y <- mnthData$Flow
x <- mnthData$Lag1
fitDEC <- locfit(y~x, data=mnthData, alpha=0.3, deg=1)
resDEC <- residuals(fitDEC)
plot(fitDEC, get.data=T, xlab="mnth Yi-1 flow", ylab="mnth Yi flow")
```

