---
title: "K-NN Stochastic Simulation"
output: html_notebook
---

```{r}
# split by months
splitMnth <- split(gsNF95, gsNF95$Month)
```

### Bootstrap K nearest neighbors streamflow algorithm

1. fit local polynomial model
2. save residuals
3. estimate mean value for y_t 
4. find k-nearest neighbors of y_t-1
5. assign a distance-based weight to neighbors 
6. normalize weight > metric
7. use metric to sample a residual from the neighbor residuals
8. calculate next month flow estimate
9. repeat for other months and successive periods to create flow sequences

```{r}
# set mean September flow as initial value
avgSEP <- mnthMean$mnthMean[10]
```

Simulate water year streamflow time series: 90 years/1080 months
```{r}
simulateWY <- function(initSep) {
  sim90WY <- list()
  simEstMuOct <- list()
  for (i in 1:90){

    # OCT ##############################################################

    estMuOCT <- predict(fitOCT, initSep) 
    estMuOCT
    
    # get Lag1 k-nearest-neighbors
    mnthData <- splitMnth$`10` 
    y <- mnthData$Flow[-1] # [-1,] trim for OCT
    x <- mnthData$Lag1[-1] # [-1,] trim for OCT
    muMnthLag <- initSep
    knnMnthLag <- get.knnx(data=x, query=muMnthLag, k=10, algorithm=c("kd_tree", "cover_tree", "CR", "brute"))

    # get Lag1 k-nearest-neighbor residuals
    knnResMnth <- resOCT[knnMnthLag$nn.index]

    # normalized nearest-neighbor distance weight vector
    sumDist <- sum(knnMnthLag$nn.dist)
    wtvec <- knnMnthLag$nn.dist/sumDist

    # simulated next time step value: y*t = yˆ*t + e*t (estimated mean yˆ*t + sampled residual e*t)
    simOCT <- estMuOCT + sample(knnResMnth,1,prob=wtvec)
    
    # NOV ##############################################################

    estMuNOV <- predict(fitNOV, simOCT)
    
    # get Lag1 k-nearest-neighbors
    mnthData <- splitMnth$`11` 
    y <- mnthData$Flow
    x <- mnthData$Lag1 
    muMnthLag <- simOCT 
    knnMnthLag <- get.knnx(data=x, query=muMnthLag, k=10, algorithm=c("kd_tree", "cover_tree", "CR", "brute"))
    
    # get Lag1 k-nearest-neighbor residuals
    knnResMnth <- resNOV[knnMnthLag$nn.index]  
    
    # normalized nearest-neighbor distance weight vector
    sumDist <- sum(knnMnthLag$nn.dist)
    wtvec <- knnMnthLag$nn.dist/sumDist

    # simulated next time step value
    simNOV <- estMuNOV + sample(knnResMnth,1,prob=wtvec)
    
    # DEC ##############################################################

    estMuDEC <- predict(fitDEC, simNOV) 
    
    # get Lag1 k-nearest-neighbors
    mnthData <- splitMnth$`12` 
    y <- mnthData$Flow # for graph
    x <- mnthData$Lag1 
    muMnthLag <- simNOV 
    knnMnthLag <- get.knnx(data=x, query=muMnthLag, k=10, algorithm=c("kd_tree", "cover_tree", "CR", "brute"))
    
    # get Lag1 k-nearest-neighbor residuals
    knnResMnth <- resDEC[knnMnthLag$nn.index] 
    
    # normalized nearest-neighbor distance weight vector
    sumDist <- sum(knnMnthLag$nn.dist)
    wtvec <- knnMnthLag$nn.dist/sumDist
    
    # simulated next time step value
    simDEC <- estMuDEC + sample(knnResMnth,1,prob=wtvec)  
    
    # JAN ##############################################################

    estMuJAN <- predict(fitJAN, simDEC)  

    # get Lag1 k-nearest-neighbors 
    mnthData <- splitMnth$`01`  
    y <- mnthData$Flow # for graph
    x <- mnthData$Lag1 
    muMnthLag <- simDEC  
    knnMnthLag <- get.knnx(data=x, query=muMnthLag, k=10, algorithm=c("kd_tree", "cover_tree", "CR", "brute"))
    
    # get Lag1 k-nearest-neighbor residuals
    knnResMnth <- resJAN[knnMnthLag$nn.index] 
    
    # normalized nearest-neighbor distance weight vector
    sumDist <- sum(knnMnthLag$nn.dist)
    wtvec <- knnMnthLag$nn.dist/sumDist
    
    # simulated next time step value
    simJAN <- estMuJAN + sample(knnResMnth,1,prob=wtvec)  
    
    # FEB ##############################################################

    estMuFEB <- predict(fitFEB, simJAN) 
    
    # get Lag1 k-nearest-neighbors 
    mnthData <- splitMnth$`02`  
    y <- mnthData$Flow # for graph
    x <- mnthData$Lag1 
    muMnthLag <- simJAN 
    knnMnthLag <- get.knnx(data=x, query=muMnthLag, k=10, algorithm=c("kd_tree", "cover_tree", "CR", "brute"))
    
    # get Lag1 k-nearest-neighbor residuals
    knnResMnth <- resFEB[knnMnthLag$nn.index] 
    
    # normalized nearest-neighbor distance weight vector
    sumDist <- sum(knnMnthLag$nn.dist)
    wtvec <- knnMnthLag$nn.dist/sumDist
    
    # simulated next time step value
    simFEB <- estMuFEB + sample(knnResMnth,1,prob=wtvec) 
    
    # MAR ##############################################################

    estMuMAR <- predict(fitMAR, simFEB) 
    
    # get Lag1 k-nearest-neighbors 
    mnthData <- splitMnth$`03`  
    y <- mnthData$Flow # for graph
    x <- mnthData$Lag1 
    muMnthLag <- simFEB  
    knnMnthLag <- get.knnx(data=x, query=muMnthLag, k=10, algorithm=c("kd_tree", "cover_tree", "CR", "brute"))
    
    # get Lag1 k-nearest-neighbor residuals
    knnResMnth <- resMAR[knnMnthLag$nn.index] 
    
    # normalized nearest-neighbor distance weight vector
    sumDist <- sum(knnMnthLag$nn.dist)
    wtvec <- knnMnthLag$nn.dist/sumDist
    
    # simulated next time step value
    simMAR <- estMuMAR + sample(knnResMnth,1,prob=wtvec) 

    # APR ##############################################################

    estMuAPR <- predict(fitAPR, simMAR)  
    
    # get Lag1 k-nearest-neighbors 
    mnthData <- splitMnth$`04`  
    y <- mnthData$Flow # for graph
    x <- mnthData$Lag1 
    muMnthLag <- simMAR  
    knnMnthLag <- get.knnx(data=x, query=muMnthLag, k=10, algorithm=c("kd_tree", "cover_tree", "CR", "brute"))
    
    # get Lag1 k-nearest-neighbor residuals
    knnResMnth <- resAPR[knnMnthLag$nn.index] 
    
    # normalized nearest-neighbor distance weight vector
    sumDist <- sum(knnMnthLag$nn.dist)
    wtvec <- knnMnthLag$nn.dist/sumDist
    
    # simulated next time step value
    simAPR <- estMuAPR + sample(knnResMnth,1,prob=wtvec)  
    
    # MAY ##############################################################

    estMuMAY <- predict(fitMAY, simAPR)  

    # get Lag1 k-nearest-neighbors 
    mnthData <- splitMnth$`05`  
    y <- mnthData$Flow # for graph
    x <- mnthData$Lag1 
    muMnthLag <- simAPR  
    knnMnthLag <- get.knnx(data=x, query=muMnthLag, k=10, algorithm=c("kd_tree", "cover_tree", "CR", "brute"))
    
    # get Lag1 k-nearest-neighbor residuals
    knnResMnth <- resMAY[knnMnthLag$nn.index] 
    
    # normalized nearest-neighbor distance weight vector
    sumDist <- sum(knnMnthLag$nn.dist)
    wtvec <- knnMnthLag$nn.dist/sumDist
    
    # simulated next time step value
    simMAY <- estMuMAY + sample(knnResMnth,1,prob=wtvec)
    
    # JUN ##############################################################

    estMuJUN <- predict(fitJUN, simMAY) 
    
    # get Lag1 k-nearest-neighbors 
    mnthData <- splitMnth$`06`  
    y <- mnthData$Flow # for graph
    x <- mnthData$Lag1 
    muMnthLag <- simMAY  
    knnMnthLag <- get.knnx(data=x, query=muMnthLag, k=10, algorithm=c("kd_tree", "cover_tree", "CR", "brute"))

    # get Lag1 k-nearest-neighbor residuals
    knnResMnth <- resJUN[knnMnthLag$nn.index]
    
    # normalized nearest-neighbor distance weight vector
    sumDist <- sum(knnMnthLag$nn.dist)
    wtvec <- knnMnthLag$nn.dist/sumDist
    
    # simulated next time step value
    simJUN <- estMuJUN + sample(knnResMnth,1,prob=wtvec) 
    
    # JUL ##############################################################

    estMuJUL <- predict(fitJUL, simJUN)  
    
    # get Lag1 k-nearest-neighbors 
    mnthData <- splitMnth$`07`  
    y <- mnthData$Flow # for graph
    x <- mnthData$Lag1 
    muMnthLag <- simJUN 
    knnMnthLag <- get.knnx(data=x, query=muMnthLag, k=10, algorithm=c("kd_tree", "cover_tree", "CR", "brute"))
    
    # get Lag1 k-nearest-neighbor residuals
    knnResMnth <- resJUL[knnMnthLag$nn.index] 
    # knnResMnth
    
    # normalized nearest-neighbor distance weight vector
    sumDist <- sum(knnMnthLag$nn.dist)
    wtvec <- knnMnthLag$nn.dist/sumDist
    
    # simulated next time step value
    simJUL <- estMuJUL + sample(knnResMnth,1,prob=wtvec) 
    
    # AUG ##############################################################

    estMuAUG <- predict(fitAUG, simJUL)  

    # get Lag1 k-nearest-neighbors 
    mnthData <- splitMnth$`08`
    y <- mnthData$Flow # for graph
    x <- mnthData$Lag1 
    muMnthLag <- simJUL
    knnMnthLag <- get.knnx(data=x, query=muMnthLag, k=10, algorithm=c("kd_tree", "cover_tree", "CR", "brute"))
    
    # get Lag1 k-nearest-neighbor residuals
    knnResMnth <- resAUG[knnMnthLag$nn.index]
    # knnResMnth
    
    # normalized nearest-neighbor distance weight vector
    sumDist <- sum(knnMnthLag$nn.dist)
    wtvec <- knnMnthLag$nn.dist/sumDist
    
    # simulated next time step value
    simAUG <- estMuAUG + sample(knnResMnth,1,prob=wtvec) 
    
    # SEP ##############################################################

    estMuSEP <- predict(fitSEP, simAUG) 
    
    # get Lag1 k-nearest-neighbors 
    mnthData <- splitMnth$`09`
    y <- mnthData$Flow # for graph
    x <- mnthData$Lag1 
    muMnthLag <- simAUG
    knnMnthLag <- get.knnx(data=x, query=muMnthLag, k=10, algorithm=c("kd_tree", "cover_tree", "CR", "brute"))

    # get Lag1 k-nearest-neighbor residuals
    knnResMnth <- resSEP[knnMnthLag$nn.index]
    
    # normalized nearest-neighbor distance weight vector
    sumDist <- sum(knnMnthLag$nn.dist)
    wtvec <- knnMnthLag$nn.dist/sumDist
    
    # simulated next time step value
    simSEP <- estMuSEP + sample(knnResMnth,1,prob=wtvec) 
    
    simWY <- c(simOCT,simNOV,simDEC,simJAN,simFEB,simMAR,simAPR,simMAY,simJUN,simJUL,simAUG,simSEP)

    sim90WY[[i]] <- simWY
    initSep <- simWY[12]
  }
  return(sim90WY)
}
```

