---
title: "Streamflow Data Exploration"
output: html_notebook:
fig_width: 6 
fig_height: 4 
---

```{r, echo=FALSE}
library(readxl)
library(ggplot2)
library(plyr)
library(dplyr)
library(lubridate)
library(moments)
library(scales)
library(sm)
library(stats)
library(gridExtra)
```

#### Data Source: United States Geological Survey

Colorado River Streamflow for USGS Gauge 09072500 at Glenwood Springs, CO
  
Natural Streamflow for the period 1906 to 2015  
Colorado River at Glenwood Springs, CO, USGS Guage 09072500  
Source url: https://www.usbr.gov/lc/region/g4000/NaturalFlow/current.html  

Load annual natural streamflow data
```{r}
library(readxl)
gsWY <- read_xlsx("~/Documents/PROJECTS/Hydro/Data/Colorado River/NaturalFlows1906-2015.xlsx", 
                  sheet = 'AnnualWYTotal Natural Flow',
                  range = "A5:B115", col_names = TRUE)
                  #col_types = c("date","numeric")
```

Clean up annual streamflow data  
1. subset data range to water years 1906 - 1995  
2. Convert data units from acre-feet/year to m^3/sec  
```{r}
endWY <- which(gsWY$`Water Year`==1995)
gsWY95 <- gsWY[0:endWY,]
gsWY95$`(ac-ft/yr)` <- gsWY95$`(ac-ft/yr)`*0.00003919
names(gsWY95) <- c("WaterYear","Flow")
```

### Annual Streamflow Time Series Plot
```{r}
plotWY <- ggplot(gsWY95, aes(WaterYear, Flow)) + 
  geom_point(size=0.9,alpha=0.35,colour="blue") + 
  geom_line(alpha=0.5,colour="blue") +
    xlab("") + ylab(expression(flow~"("*m^{3}*"/s)")) +
    labs(title = "Monthly Natural Streamflow", subtitle = "USGS Guage 09072500, Colorado River near Glenwood Springs, CO")
plotWY
```

---

Load monthly streamflow data
```{r}
library(readxl)
gsMNF <- read_xlsx("~/Documents/PROJECTS/Hydro/Data/Colorado River/NaturalFlows1906-2015.xlsx", 
                  sheet = 'Total Natural Flow',
                  range = "A5:B1328", col_names = TRUE,
                  col_types = c("date","numeric"))
```

Clean up monthly streamflow data  
1. assign column names  
2. convert POSIXct datetime to date class variable, with format YYYY-MM-DD  
3. convert acre-feet per month to cubic meters per second  
4. subset data range to water years 1906 - 1995  
5. Convert ac-ft/m to m^3/s  
6. add lag -1 and month factor columns  

```{r}
names(gsMNF) <- c("Date","Flow")
gsMNF$Date <- as.Date(gsMNF$Date, format="%Y-%m-%d")
endMNF <- which(gsMNF$Date=="1995-09-30")
gsMNF95 <- gsMNF[0:endMNF,]
gsMNF95$Flow <- gsMNF95$Flow*0.00046936
gsMNF95$Lag1 <- lag(gsMNF95$Flow, k=1)
gsMNF95$Month <- as.factor(format(gsMNF95$Date,'%m'))
```

### Monthly Streamflow Time Series Plot  
```{r, echo=FALSE}
plotMTS <- ggplot(gsMNF95, aes(Date, Flow)) + 
  geom_point(size=0.75,alpha=0.35,colour="blue") + 
  geom_line(alpha=0.5,colour="blue") +
  theme(axis.text.x = element_text(angle = 90)) +
  (scale_x_date(breaks=date_breaks("5 year"), labels=date_format("%b %Y"), 
                limits = c(as.Date("1905-10-31"), as.Date("1995-09-30")))) +
  (scale_y_continuous(labels = scales::comma)) +
  xlab("") + ylab(expression(flow~"("*m^{3}*"/s)")) +
  labs(title = "Monthly Natural Streamflow", subtitle = "USGS Guage 09072500, Colorado River near Glenwood Springs, CO")
plotMTS
```

---

### Model Comparison Statistics  

Monthly streamflow means, standard deviation, coefficient of skewness, maximum and minimum flows
```{r, echo=FALSE}
# summarize time series monthly statistics with ddply (split-apply-combine)
mnthMean <- ddply(gsMNF95, .(Month), summarize, mnthMean=mean(Flow))
mnthSD <- ddply(gsMNF95, .(Month), summarize, mnthSD=sd(Flow))
mnthSkew <- ddply(gsMNF95, .(Month), summarize, mnthSkew=skewness(Flow))
mnthMin <- ddply(gsMNF95, .(Month), summarize, mnthMin=min(Flow))
mnthMax <- ddply(gsMNF95, .(Month), summarize, mnthMax=max(Flow))
mnthStats <- as.data.frame(matrix(nrow=12,ncol=1))
mnthStats$Mean <- round(mnthMean$mnthMean,2)
mnthStats$SD <- round(mnthSD$mnthSD,2)
mnthStats$Skew <- round(mnthSkew$mnthSkew,2)
mnthStats$Min <- round(mnthMin$mnthMin,2)
mnthStats$Max <- round(mnthMax$mnthMax,2)
mnthStats$V1 <- c("Oct","Nov","Dec","Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep")
mnthStats
```

---

### Distribution of Monthly Streamflow  

Plot data: reorder month factor levels and rename month abbreviation
```{r}
gsMNF95$MonthFac = factor(gsMNF95$Month,levels(gsMNF95$Month)[c(10:12,1:9)])
gsMNF95$MonthFac <- mapvalues(gsMNF95$Month, c("01","02","03","04","05","06","07","08","09","10","11","12"), to = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
```

```{r}
g <- ggplot(data = gsMNF95, mapping = aes(x = MonthFac, y = Flow))
g + geom_boxplot(size=0.5) + 
  geom_jitter(aes(color = Flow), size=0.6, show.legend=FALSE, width=0.25) + 
  ylab(expression(streamflow~"("*m^{3}*"/s)")) + xlab("") +
  labs(title = "Monthly Natural Streamflow, 1906 - 1995", subtitle = "USGS Guage 09072500, Colorado River near Glenwood Springs, CO")
```

### Monthly Streamflow Histograms
```{r}
d <- ggplot(data = gsMNF95, aes(x = Flow)) + geom_histogram(binwidth = 2, fill = "blue")
d + facet_wrap(~ MonthFac, ncol=4) +
labs(title = "Monthly Streamflow Histogram, 1906 - 1995", subtitle = "USGS Guage 09072500, Colorado River near Glenwood Springs, CO") + ylab("") + xlab("")
```

### Monthly Streamflow Density Estimation  

```{r}
# split monthly data
splitMnth <- split(gsMNF95, gsMNF95$Month)
```

```{r}
par(mfrow=c(2,2))

oct <- splitMnth$Oct$Flow
nov <- splitMnth$Nov$Flow
dec <- splitMnth$Dec$Flow
group.index <- rep(1:3, c(length(x), length(y), length(z)))
den.ond <- sm.density.compare(c(oct,nov,dec), group = group.index, model = "equal", col=c("blue","black","grey42"),lty=c(4,3,1), xlab="Oct Nov Dec")
legend('topright', legend=c("oct","nov","dec"), lty=c(4,3,1), col=c("blue","black","grey42"), bty='n', cex=.75)

title(outer=TRUE, line = -1, main = list("Monthly Density Comparison", cex=1.25,col="black", font=2))

jan <- splitMnth$Jan$Flow
feb <- splitMnth$Feb$Flow
mar <- splitMnth$Mar$Flow
group.index <- rep(1:3, c(length(jan), length(feb), length(mar)))
den.jfm <- sm.density.compare(c(jan,feb,mar), group = group.index, model = "equal", col=c("blue","black","grey42"),lty=c(4,3,1), xlab="Jan Feb Mar")
legend('topright', legend=c("jan","feb","mar"), lty=c(4,3,1), col=c("blue","black","grey42"), bty='n', cex=.75)

apr <- splitMnth$Apr$Flow
may <- splitMnth$May$Flow
jun <- splitMnth$Jun$Flow
group.index <- rep(1:3, c(length(apr), length(may), length(jun)))
den.amj <- sm.density.compare(c(apr,may,jun), group = group.index, model = "equal", col=c("blue","black","grey42"),lty=c(4,3,1), xlab="Apr May Jun")
legend('topright', legend=c("apr","may","jun"), lty=c(4,3,1), col=c("blue","black","grey42"), bty='n', cex=.75)

jul <- splitMnth$Jul$Flow
aug <- splitMnth$Aug$Flow
sep <- splitMnth$Sep$Flow
group.index <- rep(1:3, c(length(jul), length(aug), length(sep)))
den.jas <- sm.density.compare(c(jul,aug,sep), group = group.index, model = "equal", col=c("blue","black","grey42"),lty=c(4,3,1), xlab="Jul Aug Sep")
legend('topright', legend=c("jul","aug","sep"), lty=c(4,3,1), col=c("blue","black","grey42"), bty='n', cex=.75)
```

